### 2024-10-10 paper/paper.qmd
Starter folder added

---
title: ""
author: ""
thanks: "Code and data are available at: [GitHub repo Link]"
date: today
date-format: long
format: pdf
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(here)
library(tidyverse)
options(scipen=999)
```

```{r}
#| include: false
#| warning: false
#| message: false
radiohead  <- readRDS(here("spotify_api/data/00-raw_data/radiohead.rds"))
radiohead
```

```{r}
#| warning: false
#| message: false
#| echo:    false
radiohead <- as_tibble(radiohead)

radiohead |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (ms)"
  )
```

```{r}
#| warning: false
#| message: false
#| echo:    false
rbind( the_national, radiohead) |>
  select(artist_name, album_release_date, valence) |>
  mutate(album_release_date = ymd(album_release_date)) |> 
  ggplot(aes( x = album_release_date, y = valence, color = artist_name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(facets = vars(artist_name), dir = "v") +
  labs(
    x = "Album release date",
    y = "Valence",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

### 2024-10-10 scripts/00-download_data.R

@@ -1,0 +1,32 @@
+#### Preamble ####
+# Purpose: 
+# Author: 
+# Date: 
+# Contact: 
+# License: MIT
+
+
+#### Workspace setup ###
+
+# install.packages("devtools")
+# devtools::install_github('charlie86/spotifyr', force = TRUE)
+# install.packages("here")
+
+library(spotifyr)
+library(usethis)
+library(here)
+
+edit_r_environ()
+
+#### Download data ####
+
+radiohead <- get_artist_audio_features("radiohead")
+taylor_swift <- get_artist_audio_features("06HL4z0CvFAxyc27GXpf02")
+the_national <- get_artist_audio_features("the national")
+
+
+#### Save data ####
+
+saveRDS(radiohead, "C:/UofT 2022/STA304/TP1/spotify_api/data/00-raw_data/radiohead.rds")
+saveRDS(taylor_swift, "taylor_swift.rds")
+saveRDS(the_national, "C:/UofT 2022/STA304/TP1/spotify_api/data/00-raw_data/the_national.rds")

### 2024-10-10 scripts/01-data_visualization.R
Data Visualization with Code

@@ -1,0 +1,37 @@
+library(tidyverse)
+library(ggplot2)
+library(dplyr)
+
+the_chainsmokers <- readRDS(here("data/00-raw_data/the_chainsmokers.rds"))
+the_chainsmokers <- as_tibble(the_chainsmokers)
+
+the_chainsmokers |>
+  mutate(album_release_date = ymd(album_release_date)) |>
+  ggplot(aes(
+    x = album_release_date,
+    y = duration_ms,
+    group = album_release_date
+  )) +
+  geom_boxplot() +
+  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
+  theme_minimal() +
+  labs(
+    x = "Album release date",
+    y = "Duration of song (ms)"
+  )
+
+audio_features <- c('danceability', 'energy', 'speechiness', 'acousticness', 
+                    'instrumentalness', 'liveness', 'valence')
+album_comparison <- the_chainsmokers %>%
+  group_by(album_name) %>%
+  summarise(across(all_of(audio_features), mean, na.rm = TRUE))
+album_comparison_long <- tidyr::pivot_longer(album_comparison, 
+                                             cols = audio_features, 
+                                             names_to = "feature", 
+                                             values_to = "mean_value")
+ggplot(album_comparison_long, aes(x = album_name, y = mean_value, fill = feature)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(x = "Album Name", y = "Mean Value", 
+       title = "Comparison of Audio Features by Album (Excluding Loudness and Tempo)") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  theme_minimal()

### 2024-10-10 paper/paper.qmd
Code Added to the Paper

@@
 #| message: false
-radiohead  <- readRDS(here("spotify_api/data/00-raw_data/radiohead.rds"))
-radiohead
+the_chainsmokers <- readRDS(here("data/00-raw_data/the_chainsmokers.rds"))
+the_chainsmokers
 ```
@@
 #| echo:    false
-radiohead <- as_tibble(radiohead)
-
-radiohead |>
+the_chainsmokers |>
   mutate(album_release_date = ymd(album_release_date)) |>
@@
 #| echo:    false
-rbind( the_national, radiohead) |>
-  select(artist_name, album_release_date, valence) |>
-  mutate(album_release_date = ymd(album_release_date)) |> 
-  ggplot(aes( x = album_release_date, y = valence, color = artist_name)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth() +
-  theme_minimal() +
-  facet_wrap(facets = vars(artist_name), dir = "v") +
-  labs(
-    x = "Album release date",
-    y = "Valence",
-    color = "Artist"
-  ) +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
+#| fig-align: "center"
+#| fig-width: 10
+#| fig-height: 6
+audio_features <- c('danceability', 'energy', 'speechiness', 'acousticness', 
+                    'instrumentalness', 'liveness', 'valence')
+the_chainsmokers_filtered <- the_chainsmokers %>%
+  filter(album_name != "So Far So Good (lofi remixes)")
+
+album_comparison <- the_chainsmokers_filtered %>%
+  group_by(album_name) %>%
+  summarise(across(all_of(audio_features), mean, na.rm = TRUE))
+
+album_comparison_long <- tidyr::pivot_longer(album_comparison, 
+                                             cols = audio_features, 
+                                             names_to = "feature", 
+                                             values_to = "mean_value")
+ggplot(album_comparison_long, aes(x = album_name, y = mean_value, fill = feature)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(x = "Album Name", y = "Mean Value", 
+       title = "Comparison of Audio Features by Album (Excluding Loudness and Tempo)") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8, margin = margin(t = 10))) + 
+  theme_minimal()
 ```

### 2024-10-11 scripts/00-setup.R
updated setup script

@@ -1,0 +1,33 @@
+#### Preamble ####
+# Purpose: Set up the workspace for the project, including library loading and API credentials.
+# Author: Jerry Xia
+# Date: 10 October 2024
+# Contact: Jerry.xia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Spotify Developer Account with valid Client ID and Client Secret, saved to '.Renviron'
+# Any other information needed? Ensure access to Spotify API and necessary packages installed
+
+#### Workspace setup ####
+# Load required libraries
+# install.packages("spotifyr")
+library(here)
+library(readr)
+library(dplyr)
+library(usethis)
+library(magrittr)
+library(spotifyr)
+library(tidyverse)
+library(lubridate)
+library(kableExtra)
+
+#### Set up API credentials ####
+# Ensure your Spotify Client ID and Secret are stored in your .Renviron file
+# You can open the .Renviron file with usethis and add credentials if not already done
+
+# Spotify Client ID and Secret should be stored in .Renviron
+Sys.setenv(SPOTIFY_CLIENT_ID = Sys.getenv("SPOTIFY_CLIENT_ID"))
+Sys.setenv(SPOTIFY_CLIENT_SECRET = Sys.getenv("SPOTIFY_CLIENT_SECRET"))
+
+#### Test Spotify API connection ####
+access_token <- get_spotify_access_token()
+print(access_token)  # You should see access token 

### 2024-10-12 scripts/01-download_data.R
updated all scripts

@@ -1,32 +1,20 @@
 #### Preamble ####
-# Purpose: 
-# Author: 
-# Date: 
-# Contact: 
+# Purpose: Download data about The Chainsmokers from Spotify API and save it as .rds
+# Author: Jerry Xia
+# Date: 10 October 2024
+# Contact: Jerry.xia@mail.utoronto.ca
 # License: MIT
+# Pre-requisites: Spotify API access and credentials, '00-setup.R' executed
+# Any other information needed? 
 
+#### Download Chainsmokers Data ####
+# Use the spotifyr package to download data about The Chainsmokers
+# use Spotify artist ID to retrieve the data
 
-#### Workspace setup ###
+chainsmokers_data <- get_artist_audio_features('the chainsmokers')
 
-# install.packages("devtools")
-# devtools::install_github('charlie86/spotifyr', force = TRUE)
-# install.packages("here")
+# Save the raw data
+saveRDS(chainsmokers_data, file = here("data", "raw_data", "chainsmokers.rds"))
 
-library(spotifyr)
-library(usethis)
-library(here)
-
-edit_r_environ()
-
-#### Download data ####
-
-radiohead <- get_artist_audio_features("radiohead")
-taylor_swift <- get_artist_audio_features("06HL4z0CvFAxyc27GXpf02")
-the_national <- get_artist_audio_features("the national")
-
-
-#### Save data ####
-
-saveRDS(radiohead, "C:/UofT 2022/STA304/TP1/spotify_api/data/00-raw_data/radiohead.rds")
-saveRDS(taylor_swift, "taylor_swift.rds")
-saveRDS(the_national, "C:/UofT 2022/STA304/TP1/spotify_api/data/00-raw_data/the_national.rds")
+# Confirm that the data has been saved
+print("Raw data has been saved to data/raw_data/chainsmokers.rds")

### 2024-10-12 scripts/02-data_cleaning.R

@@ -1,37 +1,22 @@
-library(tidyverse)
-library(ggplot2)
-library(dplyr)
+#### Preamble ####
+# Purpose: Clean the downloaded data for The Chainsmokers
+# Author: Jerry Xia
+# Date: 10 October 2024
+# Contact: Jerry.xia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 'chainsmokers.rds' saved in 'data/raw_data'
 
-the_chainsmokers <- readRDS(here("data/00-raw_data/the_chainsmokers.rds"))
-the_chainsmokers <- as_tibble(the_chainsmokers)
+#### Load Raw Data ####
+raw_data <- readRDS(here("data", "raw_data", "chainsmokers.rds"))
 
-the_chainsmokers |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(
-    x = album_release_date,
-    y = duration_ms,
-    group = album_release_date
-  )) +
-  geom_boxplot() +
-  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
-  theme_minimal() +
-  labs(
-    x = "Album release date",
-    y = "Duration of song (ms)"
-  )
+#### Clean Data ####
+cleaned_data <- raw_data %>%
+  mutate(album_release_date = ymd(album_release_date)) %>%
+  select(album_name, album_release_date, danceability, energy, acousticness, valence, speechiness, tempo) %>%
+  filter(!is.na(album_release_date))
 
-audio_features <- c('danceability', 'energy', 'speechiness', 'acousticness', 
-                    'instrumentalness', 'liveness', 'valence')
-album_comparison <- the_chainsmokers %>%
-  group_by(album_name) %>%
-  summarise(across(all_of(audio_features), mean, na.rm = TRUE))
-album_comparison_long <- tidyr::pivot_longer(album_comparison, 
-                                             cols = audio_features, 
-                                             names_to = "feature", 
-                                             values_to = "mean_value")
-ggplot(album_comparison_long, aes(x = album_name, y = mean_value, fill = feature)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(x = "Album Name", y = "Mean Value", 
-       title = "Comparison of Audio Features by Album (Excluding Loudness and Tempo)") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
-  theme_minimal()
+# Save the cleaned data
[…]

### 2024-10-12 scripts/03-data_visualization.R

@@ -1,0 +1,60 @@
+#### Preamble ####
+# Purpose: Visualize key features of The Chainsmokers' data for analysis
+# Author: Jerry Xia
+# Date: 10 October 2024
+# Contact: Jerry.xia@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 'clean_chainsmokers_data.rds' saved in 'data/analysis_data'
+
+#### Load Cleaned Data ####
+clean_data <- readRDS(here("data", "analysis_data", "clean_chainsmokers_data.rds"))
+
+#### Visualizations ####
+
+# 1. Energy Over Time
+energy_plot <- clean_data %>%
+  ggplot(aes(x = album_release_date, y = energy, color = album_name)) +
+  geom_line(stat = "summary", fun = "mean", size = 1.2) +
+  geom_point(alpha = 0.5) +
+  theme_minimal() +
+  labs(
+    x = "Album Release Date",
+    y = "Energy",
+    color = "Album"
+  ) +
+  theme(legend.position = "bottom")
+
+ggsave(here("other", "figures", "energy_plot.png"), plot = energy_plot)
+
+# 2. Valence Over Time
+valence_plot <- clean_data %>%
+  ggplot(aes(x = album_release_date, y = valence, color = album_name)) +
+  geom_line(stat = "summary", fun = "mean", size = 1.2) +
+  geom_point(alpha = 0.5) +
+  theme_minimal() +
+  labs(
+    x = "Album Release Date",
+    y = "Valence",
+    color = "Album"
+  ) +
+  theme(legend.position = "bottom")
+
+ggsave(here("other", "figures", "valence_plot.png"), plot = valence_plot)
+
+# 3. Danceability vs. Acousticness Over Time
+dance_acoustic_plot <- clean_data %>%
+  ggplot(aes(x = album_release_date)) +
+  geom_line(aes(y = danceability, color = "Danceability"), stat = "summary", fun = "mean", size = 1.2) +
+  geom_line(aes(y = acousticness, color = "Acousticness"), stat = "summary", fun = "mean", size = 1.2) +
+  theme_minimal() +
+  labs(
+    x = "Album Release Date",
+    y = "Score",
+    color = "Feature"
+  ) +
+  theme(legend.position = "bottom")
+
+ggsave(here("other", "figures", "dance_acoustic_plot.png"), plot = dance_acoustic_plot)
+
+# Confirm that figures have been saved
+print("Energy, Valence, and Danceability vs Acousticness plots saved to 'other/figures'")

### 2024-10-12 paper/paper.qmd
improve readability

--- 2024-10-10 paper/paper.qmd
@@
 ---
-title: ""
-author: ""
-thanks: "Code and data are available at: [GitHub repo Link]"
-date: today
+title: "The Chainsmokers’ Musical Evolution"
+subtitle: "From Pop-Infused EDM to a Consistent Electronic Core"
+author: 
+  - name: Jerry Xia
+  - name: Peter Fan
+  - name: Jason Yang
+thanks: "Code and data are available at: <https://github.com/ycfan0991/analyzing_the_chainsmokers_streaming_data>"
+date: "October 10, 2024"
 date-format: long
+abstract: "This paper analyzes the evolution of The Chainsmokers’ music by examining key audio features such as acousticness, energy, valence, and danceability using data from the Spotify API.
+The study focuses on the band's most popular albums, showcasing their transition through various pop-influenced styles while retaining a consistent electronic production core.
+By analyzing these attributes across their discography, this study provides insights into how The Chainsmokers adapted their music while maintaining high energy and rhythmic accessibility."
 format: pdf
+toc: true
+toc-depth: 3
+number-sections: true
 bibliography: references.bib
@@
 #| message: false
+
+# Load required libraries
+library(dplyr)
+library(ggplot2)
 library(here)
-library(tidyverse)
-options(scipen=999)
+library(readr)
+library(magrittr)
+library(lubridate)
+library(spotifyr)
 ```
+
+# Introduction
+
+The Chainsmokers are a globally recognized DJ duo known for their electronic dance music (EDM) blended with pop influences.
+Since their breakout in 2014 with "Selfie," they have consistently released chart-topping singles and albums that blend infectious beats with mainstream pop appeal.
+This fusion has allowed them to carve out a unique space in the music industry, where they collaborate with various vocalists to create a signature sound that resonates across diverse audiences.
+
+This paper seeks to explore the evolution of The Chainsmokers' sound by analyzing key audio features such as **acousticness**, **energy**, **valence**, **danceability**, and **speechiness** using data sourced from the Spotify API.
+By examining these attributes across their most popular albums, this study aims to uncover trends in their musical style and provide insights into how they have adapted to changes in the music landscape while maintaining core elements of their production.
+
+Focusing on albums such as *Memories...Do Not Open*, *Sick Boy*, and *World War Joy*, this paper will analyze how The Chainsmokers have balanced pop trends with their electronic roots.
+The analysis will highlight their consistent use of high-energy tracks and their exploration of emotional tone (valence) and danceability.
+
+This paper is structured as follows: @sec-data presents the data sources, methodologies, and visualization of the results.
+The findings are then discussed in @sec-dis, and additional detailed information is provided in the appendix at @sec-app.
+
+# Data {#sec-data}
+
+This project is guided by Rohan Alexander's work and his book on data analysis [@citeTbook].
+The data used in this paper were downloaded, cleaned, and analyzed using the programming language `R` [@citeR] and the following R packages: `dplyr` [@citeDplyr], `usethis` [@citeUsethis], `magrittr` [@citeMagrittr], `spotifyr` [@citeSpotifyr], `tidyverse` [@citeTidyverse], `lubridate` [@citeLubridate], `here` [@citeHere], and `readr` [@citeReadr].
+
+## Source
+
+Data for this paper was obtained from the Spotify API [@citeSpotify].
+The Spotify API provides comprehensive track-level data, including details about acoustic features, popularity metrics, and track metadata for artists.
+Using the `spotifyr` package [@citeSpotifyr], we were able to query The Chainsmokers' discography and retrieve key information such as track-level audio features and album release dates.
+
+The dataset includes all albums released by The Chainsmokers, such as *Memories...Do Not Open*, *Sick Boy*, and *World War Joy*.
+Each track in the dataset is associated with its album release date, enabling a temporal analysis of their musical evolution.
+
+## Measurement
+
+Spotify provides several audio features that are key to understanding the musical characteristics of tracks.
+These features, while not exhaustively described by Spotify, are commonly interpreted as follows:
+
+- **Acousticness**: A measure of the likelihood of a track being acoustic. Higher values indicate that the track is more acoustic and uses non-electronic instruments.
+- **Energy**: A measure of intensity and activity, capturing the dynamic and energetic feel of a track. High energy typically represents fast-paced, intense music.
+- **Valence**: A measure of musical positiveness. Tracks with higher valence are considered more positive and cheerful, while lower valence suggests a more somber or melancholic tone.
+- **Danceability**: A measure of how suitable a track is for dancing, based on its rhythm stability, tempo, and beat strength.
+- **Speechiness**: Speechiness quantifies the presence of spoken words in a track. Higher values indicate a greater presence of speech, as found in rap or spoken word tracks.
+
+Further details about the data processing and cleaning steps can be found in the appendix at @sec-app.
 
@@
 #| message: false
-radiohead  <- readRDS(here("spotify_api/data/00-raw_data/radiohead.rds"))
-radiohead
+
+# Load the cleaned data
+clean_data <- readRDS(here::here("data", "analysis_data", "clean_chainsmokers_data.rds"))
+
+# Generate the energy plot
+energy_plot <- clean_data %>%
+  ggplot(aes(x = album_release_date, y = energy, color = album_name)) +
+  geom_line(stat = "summary", fun = "mean", size = 1.2) +
+  geom_point(alpha = 0.5) +
+  theme_minimal() +
+  labs(
+    x = "Album Release Date",
+    y = "Energy",
+    color = "Album"
+  ) +
+  theme(legend.position = "bottom")
+
+# Generate the valence plot
+valence_plot <- clean_data %>%
+  ggplot(aes(x = album_release_date, y = valence, color = album_name)) +
+  geom_line(stat = "summary", fun = "mean", size = 1.2) +
+  geom_point(alpha = 0.5) +
+  theme_minimal() +
+  labs(
+    x = "Album Release Date",
+    y = "Valence",
+    color = "Album"
+  ) +
+  theme(legend.position = "bottom")
+
+# Generate the danceability and acousticness comparison plot
+dance_acoustic_plot <- clean_data %>%
+  ggplot(aes(x = album_release_date)) +
+  geom_line(aes(y = danceability, color = "Danceability"), stat = "summary", fun = "mean", size = 1.2) +
+  geom_line(aes(y = acousticness, color = "Acousticness"), stat = "summary", fun = "mean", size = 1.2) +
+  theme_minimal() +
+  labs(
+    x = "Album Release Date",
+    y = "Score",
+    color = "Feature"
+  ) +
+  theme(legend.position = "bottom")
+```
+
+## Results
+
+This section presents the analysis of key audio features from The Chainsmokers' albums, focusing on **energy**, **valence**, **danceability**, and **acousticness**.
+The data visualizations illustrate how these features have evolved over time and provide insight into the musical trends that define The Chainsmokers' sound.
+
+### Energy Across Albums
+
+Energy is a crucial feature of The Chainsmokers' music, reflecting the intensity and dynamic nature of their tracks.
+As shown in @fig-energy, The Chainsmokers have consistently maintained high energy levels across their albums.
+Tracks from albums such as *Memories...Do Not Open* and *World War Joy* exhibit energy values that typically remain above 0.7, indicating the fast-paced, upbeat nature of their music.
+This trend is reflective of their electronic dance music (EDM) roots, where high energy is a signature element.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-energy
+#| fig-cap: "Energy Levels of The Chainsmokers' Tracks Over Time (2014-2023)"
+#| warning: false
+#| message: false
+#| fig-height: 6
+#| fig-width: 7
+
+energy_plot
+```
+@fig-energy illustrates the consistency in energy across The Chainsmokers' discography, reinforcing their reputation for creating music with high intensity and dynamic beats.
+
+Valence measures the emotional positivity of a track.
+A higher valence score indicates a more positive, cheerful sound, while lower scores suggest a more somber or melancholic tone.
+As shown in @fig-valence, there is a slight downward trend in valence in The Chainsmokers' music, particularly in their more recent albums like Sick Boy and World War Joy.
+This shift may reflect a growing maturity in their themes, moving from upbeat, pop-infused tracks to more emotionally complex compositions.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-valence
+#| fig-cap: "Valence of The Chainsmokers' Tracks Across Albums"
+#| warning: false
+#| message: false
+#| fig-height: 6
+#| fig-width: 7
+
+valence_plot
+```
+@fig-valence highlights how The Chainsmokers' emotional tone has shifted slightly over time, suggesting a nuanced evolution in their thematic choices.
+
+The Chainsmokers' music is typically characterized by its suitability for dancing, with danceability being a key feature in their tracks.
+As shown in @fig-dance_acoustic, their tracks maintain high danceability scores, with values above 0.6 across most albums.
+This reflects their commitment to creating music that is rhythmically engaging and suited for dance floors.
+
+On the other hand, acousticness remains low across their discography, indicating a consistent use of electronic production and minimal reliance on acoustic instruments.
+This reinforces their electronic music identity, as seen in albums like Sick Boy and Memories...Do Not Open, where acousticness scores remain close to 0.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-dance_acoustic
+#| fig-cap: "Danceability and Acousticness of The Chainsmokers' Tracks Over Time"
+#| warning: false
+#| message: false
+#| fig-height: 6
+#| fig-width: 7
+
+dance_acoustic_plot
+```
+@fig-dance_acoustic shows that while The Chainsmokers' music is highly danceable, their reliance on electronic elements remains central, with low acousticness scores across their albums.
+
+# Discussion {#sec-dis}
+
+The analysis of The Chainsmokers' discography reveals several key trends in their musical evolution.
+Consistent with their identity as electronic dance music (EDM) producers, The Chainsmokers maintain high **energy** levels across their tracks.
+As seen in @fig-energy, the energetic intensity of their music is a defining characteristic, particularly in albums like *Memories...Do Not Open* and *World War Joy*.
+This sustained high energy supports their reputation for creating tracks suited for both mainstream pop audiences and electronic music festivals.
+
+In contrast, the **valence** analysis shows a slight downward trend in recent albums, such as *Sick Boy* and *World War Joy*, suggesting a shift toward more complex and potentially melancholic emotional tones (@fig-valence).
+This evolution might reflect broader trends in popular music, where artists often shift from bright, upbeat themes to more introspective and emotionally nuanced topics over time.
+The Chainsmokers have managed to balance this shift in tone while still maintaining their electronic roots.
+
+The comparison of **danceability** and **acousticness** further emphasizes their electronic foundation.
+As demonstrated in @fig-dance_acoustic, The Chainsmokers consistently produce tracks with high danceability, ensuring their music remains accessible to club and festival settings.
+At the same time, **acousticness** remains low, indicating their continued reliance on electronic instruments and production techniques rather than incorporating acoustic elements.
+This distinction between danceability and acousticness highlights their ability to stay true to their EDM origins while adapting to trends in popular music.
+
+Overall, these findings suggest that while The Chainsmokers have experimented with emotional tone (valence), they have largely preserved the key elements of their signature sound—high energy, danceable beats, and minimal acoustic production.
+These elements have helped them remain relevant in a constantly evolving music industry.
+
+## Weaknesses and Next Steps
+
+While this analysis provides valuable insights into The Chainsmokers' musical evolution, there are several limitations that could be addressed in future studies.
+One notable weakness is the reliance on a relatively small set of audio features (energy, valence, danceability, acousticness, and speechiness).
+Although these features offer a useful framework for understanding their music, they do not fully capture other important aspects such as **lyrical content**, **instrumentation intricacies**, and **production techniques**.
+Future research could expand the analysis by incorporating sentiment analysis or lyrical theme extraction to assess how their thematic content has evolved alongside their sound.
+
+Another limitation is the focus on **studio-recorded tracks**.
+While these tracks represent the core of The Chainsmokers' discography, they do not account for **live performances** or **remixes**, which often play a crucial role in electronic dance music culture.
+Including data from live recordings or remix albums could provide a more holistic understanding of their artistic expression and how their tracks are adapted for different contexts.
+
+An interesting observation in the data is the **consistency in tempo** across their albums, which was not a primary focus of this study.
+Analyzing tempo could offer additional insights into their pacing and how they maintain listener engagement.
+Similarly, **loudness** could be another feature worth investigating, especially in relation to how their tracks are mastered for different platforms (e.g., streaming services vs.
+live performances).
+These additional features could help to paint a more complete picture of their production choices.
+
+Finally, **collaborations with featured artists** could be explored in more depth.
+The Chainsmokers frequently collaborate with other artists, which can significantly influence the sound and direction of individual tracks.
+An analysis of how **collaborations** impact key features like **valence** or **energy** could reveal whether these partnerships have contributed to the emotional shifts observed in their more recent works.
+
+In future work, expanding the dataset to include these additional dimensions—lyrics, live performances, tempo, loudness, and collaborations—could yield a more nuanced understanding of The Chainsmokers' music.
+This would provide a richer context for their evolution as artists and offer new perspectives on how they continue to navigate and influence the ever-changing music industry.
+
+\appendix
+
+# Appendix {#sec-app}
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-raw
+#| tbl-cap: "Raw Chainsmokers Spotify Data"
+#| warning: false
+#| message: false
+
+# Select and round key columns from the raw data
+selected_raw_data <- raw_data[, c("album_id", "danceability", "energy", "acousticness", "valence", "speechiness")]
+selected_raw_data[, 2:6] <- round(selected_raw_data[, 2:6], 2)
+
+# Display the first six rows of raw data
+kbl(selected_raw_data[1:6, ]) %>%
+  kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed"))
 ```
@@
 ```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-clean
+#| tbl-cap: "Processed Chainsmokers Spotify Data"
 #| warning: false
 #| message: false
-#| echo:    false
-radiohead <- as_tibble(radiohead)
 
-radiohead |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(
-    x = album_release_date,
-    y = duration_ms,
-    group = album_release_date
-  )) +
-  geom_boxplot() +
-  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
-  theme_minimal() +
-  labs(
-    x = "Album release date",
-    y = "Duration of song (ms)"
-  )
+# Select and round key columns from the cleaned data
+selected_clean_data <- clean_data[, c("album_id", "danceability", "energy", "acousticness", "valence", "speechiness")]
+selected_clean_data[, 2:6] <- round(selected_clean_data[, 2:6], 2)
+
+# Display the first six rows of cleaned data
+kbl(selected_clean_data[1:6, ]) %>%
+  kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed"))
 ```
+Data cleaning involved selecting the relevant audio features—danceability, energy, acousticness, valence, and speechiness—from the Spotify API data.
+Inconsistent album release dates were corrected using lubridate::ymd(), and rows with missing or invalid values were removed.
 
-```{r}
-#| warning: false
-#| message: false
-#| echo:    false
-rbind( the_national, radiohead) |>
-  select(artist_name, album_release_date, valence) |>
-  mutate(album_release_date = ymd(album_release_date)) |> 
-  ggplot(aes( x = album_release_date, y = valence, color = artist_name)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth() +
-  theme_minimal() +
-  facet_wrap(facets = vars(artist_name), dir = "v") +
-  labs(
-    x = "Album release date",
-    y = "Valence",
-    color = "Artist"
-  ) +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
-```
+The cleaned dataset (@tbl-clean) ensures consistency and completeness, as seen by comparing it to the raw data (@tbl-raw).

### 2024-10-12 paper/paper.qmd
updated files for improved readibility

@@
 library(spotifyr)
+library(kableExtra)
+
+# Load the raw data
+raw_data <- readRDS(here::here("data", "raw_data", "chainsmokers.rds"))
+clean_data <- readRDS(here::here("data", "analysis_data", "clean_chainsmokers_data.rds"))
 ```
+
+\newpage
 
@@
 Valence measures the emotional positivity of a track.
-A higher valence score indicates a more positive, cheerful sound, while lower scores suggest a more somber or melancholic tone.
-As shown in @fig-valence, there is a slight downward trend in valence in The Chainsmokers' music, particularly in their more recent albums like Sick Boy and World War Joy.
-This shift may reflect a growing maturity in their themes, moving from upbeat, pop-infused tracks to more emotionally complex compositions.
+A higher valence score suggests a more cheerful, upbeat tone, while lower scores indicate a more somber or melancholic mood.
+As shown in @fig-valence, there is a slight upward trend in valence in The Chainsmokers' music, especially in more recent albums like *So Far So Good* and *Summertime Friends*.
+This gradual increase may reflect a subtle shift toward more positive or varied emotional tones in their music, signaling an evolution in their thematic and production choices.
 
@@
 ```
-@fig-valence highlights how The Chainsmokers' emotional tone has shifted slightly over time, suggesting a nuanced evolution in their thematic choices.
+@fig-valence illustrates how The Chainsmokers’ emotional tone has evolved slightly, showcasing a more positive shift over time across their discography.
 
@@
 
+\newpage
+
 \appendix
@@
 # Appendix {#sec-app}
+
+## Raw Data Processing
+The raw data for this analysis was sourced from the Spotify API using the spotifyr [@citeSpotifyr] package.
+The initial data includes audio features for The Chainsmokers' tracks, such as danceability, energy, acousticness, valence, speechiness, and tempo.
+The cleaning process focused on ensuring temporal accuracy by filtering and formatting the album release dates.
 
@@
 #| label: tbl-raw
-#| tbl-cap: "Raw Chainsmokers Spotify Data"
-#| warning: false
-#| message: false
+#| tbl-cap: "Raw The Chainsmokers Spotify Data"
 
-# Select and round key columns from the raw data
-selected_raw_data <- raw_data[, c("album_id", "danceability", "energy", "acousticness", "valence", "speechiness")]
-selected_raw_data[, 2:6] <- round(selected_raw_data[, 2:6], 2)
+selected_raw_data <- raw_data[, c("album_name", "album_release_date", "danceability", "energy", "acousticness", "valence")]
+selected_raw_data[, 3:6] <- round(selected_raw_data[, 3:6], 2)
 
-# Display the first six rows of raw data
 kbl(selected_raw_data[1:6, ]) %>%
   kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed"))
+
 ```
+## Cleaned Data
+To clean the data, tracks missing release dates were removed, and the `album_release_date` was formatted into a consistent year-month-day format using `lubridate::ymd()`.
+Tracks with missing or invalid audio features were also excluded to maintain data integrity.
+The cleaned data was saved for further analysis.
 
@@
 #| label: tbl-clean
-#| tbl-cap: "Processed Chainsmokers Spotify Data"
-#| warning: false
-#| message: false
+#| tbl-cap: "Processed The Chainsmokers Spotify Data"
 
-# Select and round key columns from the cleaned data
-selected_clean_data <- clean_data[, c("album_id", "danceability", "energy", "acousticness", "valence", "speechiness")]
-selected_clean_data[, 2:6] <- round(selected_clean_data[, 2:6], 2)
+selected_clean_data <- clean_data[, c("album_name", "album_release_date", "danceability", "energy", "acousticness", "valence")]
+selected_clean_data[, 3:6] <- round(selected_clean_data[, 3:6], 2)
 
-# Display the first six rows of cleaned data
 kbl(selected_clean_data[1:6, ]) %>%
   kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed"))
+
 ```
-Data cleaning involved selecting the relevant audio features—danceability, energy, acousticness, valence, and speechiness—from the Spotify API data.
-Inconsistent album release dates were corrected using lubridate::ymd(), and rows with missing or invalid values were removed.
+## Data Processing Overview
+The cleaning process began by loading the raw data from a `.rds` file.
+To ensure that all tracks included in the analysis were associated with valid albums, any tracks with missing release dates were excluded.
+The date formatting was standardized using `lubridate::ymd()`, ensuring consistency across all entries.
+Additionally, audio features critical to the analysis, such as danceability, energy, acousticness, and valence, were retained while others were excluded for relevance.
+Finally, the cleaned data was saved in the `analysis_data` folder.
 
-The cleaned dataset (@tbl-clean) ensures consistency and completeness, as seen by comparing it to the raw data (@tbl-raw).
+The difference between the raw and cleaned datasets can be observed by comparing @tbl-raw and @tbl-clean.
+The final dataset contains only complete and reliable records, ensuring that subsequent analyses are based on valid data.
+
+\newpage
+
+# References {#sec-references}

### 2024-10-12 paper/paper.qmd
fixed table issue

@@
 As shown in @fig-energy, The Chainsmokers have consistently maintained high energy levels across their albums.
-Tracks from albums such as *Memories...Do Not Open* and *World War Joy* exhibit energy values that typically remain above 0.7, indicating the fast-paced, upbeat nature of their music.
+Tracks from albums such as *Memories...Do Not Open* and *World War Joy* exhibit energy values that typically remain above 0.5, indicating the fast-paced, upbeat nature of their music.
 This trend is reflective of their electronic dance music (EDM) roots, where high energy is a signature element.
@@
 
-selected_raw_data <- raw_data[, c("album_name", "album_release_date", "danceability", "energy", "acousticness", "valence")]
-selected_raw_data[, 3:6] <- round(selected_raw_data[, 3:6], 2)
+selected_raw_data <- raw_data[, c("album_id", "danceability", "energy", "acousticness", "valence")]
+selected_raw_data[, 2:5] <- round(selected_raw_data[, 2:5], 2)
 

### 2024-10-12 paper/paper.qmd
fixed italics

@@
 On the other hand, acousticness remains low across their discography, indicating a consistent use of electronic production and minimal reliance on acoustic instruments.
-This reinforces their electronic music identity, as seen in albums like Sick Boy and Memories...Do Not Open, where acousticness scores remain close to 0.
+This reinforces their electronic music identity, as seen in albums like *Sick Boy* and *Memories...Do Not Open*, where acousticness scores remain close to 0.
 
